---
title: "05.1_bkmr_si"
author: "Jagadeesh Puvvula"
date: "2025-02-14"
output: pdf_document
---

```{r}
dat_si <- read_csv(paste0(data, "basc_imputed_nov_2024.csv")) 

dat<- dat_si |>
  select(c(Pb, Hg, DMA, DDE, PBDE_47, PCB_118, PCB_138, PCB_153, PCB_180,
          PFHxS, PFOA, PFOS, BCEtP, BDCPP, DBuP, DPhP, TCS, BPA, MBP, MBZP,
          MCPP, sigma_DEHP, MEP, MIBP, di_Ethyl_OP, di_Methyl_OP, 
          B_PB, M_PB, P_PB, cotinine, home_score_total, mom_age, 
          cohort, city, sex, race_bin,  mom_edu_cat, parity_n, mat_dep_z,
          basc_ext, basc_inz, basc_bsi, brief_wm, brief_po)) |>
  rename(BCEP=BCEtP, BDCIPP=BDCPP, DNBP=DBuP, DPhP=DPhP)|>
  mutate(across(all_of(c("cohort", "city", "sex", "race_bin", 
                         "parity_n", "mom_edu_cat")), as.integer),
         log_cotinine = log2(cotinine+ 1e-6)) |>
  filter(basc_ext != 0 & basc_inz != 0 & basc_bsi != 0 & brief_wm != 0 & brief_po != 0)


```

#for 29 chemicals
```{r}
chemicals_29<- c("Pb","Hg", "DMA",
              "DDE", "PBDE_47","PCB_118","PCB_138", "PCB_153","PCB_180",
              "PFHxS", "PFOA", "PFOS",
              "BCEP", "BDCIPP", "DNBP", "DPhP",
              "TCS", "BPA",
              "di_Ethyl_OP", "di_Methyl_OP",
              "MBP","MBZP","MCPP", "sigma_DEHP","MEP","MIBP",
              "B_PB", "M_PB", "P_PB")
chem_group_29<- c(rep(1,times=3), rep(2,times=6), rep(3, times=3),
               rep(4,times=4), rep(5,times=2), rep(6,times=2),
               rep(7,times=6), rep(8,times=3))
outcomes<-c("basc_ext","basc_inz", "basc_bsi", "brief_wm", "brief_po")
covariates<- c("race_bin", "log_cotinine", "mom_edu_cat", 
               "home_score_total", "parity_n", "mom_age", "mat_dep_z")

# for all chemicals select 1:32
dat <- dat |>
  mutate_at(vars(1:29), ~log2(.+0.0000001)) |>
  drop_na() |>
  mutate_all(as.numeric)
```

=================================================================================
#BKMR loop
#Non-linearlity modeled using #knots = (15/sample size)*100
#convergence tested at 100 iterations
# TUNE nd value in knots100 object and iter in kmbayes function

#Run model (select either sex or cohort to TRUE: Convergence issue)
#for all 29 chemicals
```{r}
bkmr_sex(data=dat, 
          folder_path = "E:/BBK17/pj/basc_brief_analy/result/bkmr_si", 
          ln_mixture_vars= chemicals_29,
          groups=chem_group_29,
          outcome_vars= outcomes, 
          covariate_vars= covariates, 
          include_sex = TRUE, iter = 50000)
```

#Extracting PIPs and overall estimates
```{r}
results <- extract_data_from_folder("E:/BBK17/pj/basc_brief_analy/result/bkmr_si",
                                    select_iterations= seq(25000, 50000, by = 50),
                                    estimate_sequence= seq(0.10, 0.90, by = 0.05))
```

#prepare data for plotting
```{r}
overall_risk<- results[2] |>as.data.frame() |>
  clean_names()|>
  rename("outcome" = "overall_risk_df_measure_name",
         "cohort" = "overall_risk_df_outcome_name",
         "gender"= "overall_risk_df_data_name") |>
  mutate(cohort = fct_recode(as.factor(cohort), Pooled = "all")) |>
  mutate(outcome = fct_recode(as.factor(outcome), BASC_BSI = "bsi", BASC_EXT = "ext", BASC_INZ = "inz", BRIEF_PO = "po", BRIEF_WM = "wm")) |>
  mutate(gender = fct_recode(as.factor(gender), All = "all", Female = "female", Male = "male"))
```

#plotting overall estimates
```{r}
ggplot(overall_risk, aes(overall_risk_df_quantile, overall_risk_df_est, 
                         ymin = overall_risk_df_est - 1.96*overall_risk_df_sd, 
                         ymax = overall_risk_df_est + 1.96*overall_risk_df_sd)) + 
    geom_hline(yintercept = 00, linetype = "dashed", color = "gray") + 
    geom_pointrange(size = 0.15, aes(color = ifelse((overall_risk_df_est - 1.96 * overall_risk_df_sd) <= 0 &
                                                  (overall_risk_df_est + 1.96 * overall_risk_df_sd) >= 0,
                                                "Credible intervals with null", "Credible intervals without null"))) + 
  scale_color_manual(values = c("Credible intervals with null" = "black", "Credible intervals without null" = "#D55E00")) +
    labs(x=NULL,
       y=expression(paste(beta[italic(bkmr)]," [95% CrI]")),
       title = "",
       caption = "")+
  theme_bw()+
    facet_grid(gender~outcome, scales = "free")+
    theme(axis.line = element_line(colour = "black"),
          axis.text=element_text(size=10), 
          axis.title=element_text(size=10,face="bold"),
          strip.text = element_text(size=10),
          axis.text.x = element_text(angle = 0, hjust = 1, vjust = 1, size = 10),
          axis.text.y = element_text(size = 10), 
          panel.spacing.x=unit(0.5, "lines"),
          panel.spacing.y=unit(0.5, "lines"),
          panel.grid.minor = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.major.y = element_blank(),
          plot.caption = element_text(size = 6),
          plot.title=element_text(size = 10, hjust = 0),
          panel.border = element_blank(),
          text=element_text(size=10),
          axis.ticks.x=element_blank(),
          strip.text.x=element_text(size=10,angle=0),
          strip.text.y = element_text(size=10),
          strip.placement = "outside",
          strip.background = element_rect(fill =  "transparent"),
          legend.box.just = "center",
          legend.spacing = unit(0, "cm"),
          legend.position = "bottom",
          legend.box = "horizontal",
          legend.direction = "horizontal",
          legend.text = element_text(size = 8),
          legend.key = element_blank(),
          legend.key.height = unit(3, "mm")) +
  guides(col=guide_legend(title = " "))+
  scale_x_continuous(breaks = seq(0.2, 0.8, by = 0.3),
                     labels = seq(0.2, 0.8, by = 0.3))

```

#save effect estimate plots
```{r}
ggsave("E:/BBK17/pj/basc_brief_analy/result/bkmr_si/bkmr_29_chem.tiff", 
       width = 10,height = 6,
       dpi=300)
```

#risks for a biomarker by holding all other biomarkers at 50th percentile
```{r}
rda_files <- list.files("E:/BBK17/pj/basc_brief_analy/result/bkmr_si/", "\\.rda$", full.names = TRUE)

for (file in rda_files) {
  name <- tools::file_path_sans_ext(basename(file)) # Extract filename without extension
  assign(name, get(load(file)), envir = .GlobalEnv) # Load and assign to filename
}


# Get all objects in the global environment that match the pattern
fit_objects <- ls(pattern = "^bkmr_")

# Apply the function and store results in a dataframe
results_df <- do.call(rbind, lapply(fit_objects, function(fit_name) {
  res <- bkmr::SingVarRiskSummaries(
    fit = get(fit_name),
    qs.diff = c(0.25, 0.75),
    q.fixed = c(0.50),
    method = "exact"
  )
  res$fit_name <- fit_name # Add a column to track source object
  res
}))

# Convert to tibble for better readability
results_df <- tibble::as_tibble(results_df)

combined_df<- results_df |>
  mutate(
    gender = word(fit_name, 2, sep = "_") |> str_to_title(),
    outcome = str_extract(fit_name, "[^_]+_[^_]+$") |> toupper()
  )

write_csv(combined_df, "E:/BBK17/pj/basc_brief_analy/result/bkmr_si/risk_from_pct_25_75.csv")
```

```{r}
ggplot(combined_df, 
       aes(fct_rev(variable), est, 
           ymin = est - 1.96*sd, 
           ymax = est + 1.96*sd,
           color= est<0)) + 
  geom_hline(aes(yintercept = 0), linetype="dashed", color= "grey50") +
  scale_color_manual(values = c("TRUE" = "royalblue", "FALSE" = "black"),
                     name= "Estimate trend",
                     labels= c("Positive", "Negative")) +
  labs(y = "Estimate", x = "Biomarker")+
  geom_pointrange(position = position_dodge(width = 0.75)) + 
  coord_flip()+
  facet_grid(.~outcome+gender, scales = "free_x")+
  theme_bw()+
  theme(
    axis.title = element_text(color = "black", size = 11),
    axis.text = element_text(color = "black", size = 11),
    axis.text.x = element_text(angle = 90),
    plot.title = element_text(color = "black", size = 11),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 12, face = "bold"),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.title = element_text(size = 11, color = "black"),
    legend.text = element_text(size = 11, color = "black")
  )
```

```{r}
ggsave("E:/BBK17/pj/basc_brief_analy/result/bkmr_si/50th_pct_plot.tiff", 
       bg="white",
       width = 16,height = 8,
       dpi=300)
```


